Motion planning for assessing accessibility of protein tunnels

### 2017-09-08 paper.tex
added images (not in repo yet)
conformation selection

Therefore, we propose to measure the distance as the smallest 3D Euclidean distance between atoms placed at $\qrand$ and a configuration
$q$ being tested.
This metric is denoted $\da(\cdot,\cdot)$ on line~\ref{alg::expand:a} in Alg.~\ref{alg::expand}.
This metric supports retraction of the ligand towards $\qrand$.
As the expansion routine attempts to find such sample that minimizes this distance to $\qrand$, it supports retraction of the ligand towards $\qrand$.

@@

@@

@@

@@

@@

From motion planning point of view, both types of conformations are useful.
The `longer' conformations can better fit in a narrow passage, but their rotation inside the narrow passage is problematic.
On the other hand, the `packed' conformations cannot fit the narrow passage, but they can more likely rotate.

In the third experiment, we tested how the selection of conformations influences the ability of the planners to find a solution.
Two strategies were used to select the conformations.
In the first strategy, denoted as Packed, $n$ `packed' conformations are selected from the available set of conformations.
The second strategy, denoted as Various, selects $n$ conformations as follows.
For each conformation, the geometric center is calculated and the deviation of the distances of the atoms from the center is calculated.
The conformations are sorted according to this deviation and then $n$ conformations are selected from this sorted list uniformly (i.e., the
        conformation number $1$, then $1+n$, etc.).

>>>>

Therefore, we propose to measure the distance as the smallest 3D Euclidean distance between atoms placed at $\qrand$ and a configuration
$q$ being tested.
This metric is denoted $\da(\cdot,\cdot)$ on line~\ref{alg::expand:a} in Alg.~\ref{alg::expand}.
As the expansion routine attempts to find such sample that minimizes this distance to $\qrand$, it supports retraction of the ligand towards $\qrand$.
The fast evaluation of this metric with complexity $O(\log n)$, where $n$ is the number of atoms of the ligand, can be achieved using KD-trees.
For each conformation $\l \in L$ of the ligand, a KD-tree is build for the ligand at configuration $q=(0,0,0,0,0,0,l,\smax)$
(i.e., for ligand with no translation and no rotation).
To find a nearest atom to $\qrand$ for a ligand at the configuration $q=(x,y,z,r_x,r_y,r_z,l,s)$, the translation vector $T=(x,y,z)$ and
rotation matrix $R$ given by $r_x$, $r_y$ and $r_z$ are computed.
Then, $\qrand$ is translated to $\qrand' = R^{-1}(\qrand-T)$ and the KD-tree search is used to find nearest point towards $\qrand'$, which
is also the nearest neighbor of the ligand at $q$ to $\qrand$.

@@

\subsection{Selection of conformations for motion planing}
\label{sec::strat}

From motion planning point of view, both types of conformations are useful.
The `longer' conformations can better fit in a narrow passage, while the `packed' conformations can be better rotated at place.
To select the suitable conformations, a simple technique is proposed.  
For each conformation, the geometric center is calculated and the deviation of the distances of the atoms from the center is calculated.
The conformations are sorted according to this deviation and then the desired number of conformations are selected from this sorted list uniformly.

@@

TODO Bara: prepsat/motivace/jak se visuazlizuje

@@

{\def\a{270}
\begin{figure}
\centering
\renewcommand{\tabcolsep}{0pt}
\begin{tabular}{cccc}
\includegraphics[width=0.15\textwidth,angle=\a]{fig/m001-oneAtom-cdCount} &
\includegraphics[width=0.15\textwidth,angle=\a]{fig/m001-oneAtom-rrt_path_time} &
\includegraphics[width=0.15\textwidth,angle=\a]{fig/m037t-cdCount} & 
\includegraphics[width=0.15\textwidth,angle=\a]{fig/m037t-rrt_path_time} \\
m01 & m01 & m37 & m37 \\

\end{tabular}
\end{figure}
}

@@

The aim of this experiment is to verify the procedure to select ligand conformations for the motion planning.
The experiment was performed with the ligand DCP, for which $1300$ conformations where available and for the ligand
M040 with $7560$ available conformations.
Two strategies were tested: the strategy described in Sec.~\ref{sec::strat} (denoted `S') and a strategy,
that prefers smaller conformations (denoted `P').
In the second method, deviation of the atomic distance from the geometric center are computed and the conformations
with the lowest deviations are selected.
The results are summarized in Tab.~\ref{tabl::selection}, which shows the percentage number of frames that are traversable by the tested ligand.
For each tested ligand, 5 and 50 conformations from the available pool of conformations were selected using the strategies.
In the case of both tested ligands and 50 selected conformations, the `S' strategy leads to higher traversability rate than the `P' strategy.
For the ligand DCP, a higher traversability rates was achieved when 50 conformations are selected.
This experiments shows that selecting 50 conformations using the `S' strategy (Sec.~\ref{sec::strat}) leads to better results
than using less number of conformations or when preferring the packed conformations.

\begin{table}
\centering
\caption{\label{tab::selection}Traversability of the tunnels depending on conformation selection}
\small
\renewcommand{\tabcolsep}{3pt}
{\small
\input graphSel.tex
}
\end{table}

@@
